
# EDF to CAL

# load packages 

```{r}
library(sf)
library(data.table)
library(dplyr)
library(tmap)
library(grid)
library(stringr)
library(spdep)
```


## EDF to AI 

```{r}
ttm <- fread(file.path("london_ai.csv"))

ttm$from_id <- ttm$from_id%>%
  as.character(.)

summary(ttm)
```

## merge ptal data

```{r}
origin_sf <- fread(file.path("input/origin.csv"))%>%
  st_as_sf(.,coords = c("lat", "lon"), crs = 4326)%>%
  select("id","AI2015","PTAL2015","geometry")

origin_sf$id <- origin_sf$id%>%
  as.character(.)

AI_and_PTAL <-ttm%>%
  left_join(.,origin_sf,by= join_by("from_id"=="id"))%>%
  st_as_sf(crs = 4326)
```

## AI to CAL 

```{r}
valid <- AI_and_PTAL%>%
  filter(AI2015 > 0)

count_all <- length(valid$PTAL2015)

PTAL_summary <- AI_and_PTAL%>%
  group_by(PTAL2015)%>%
  summarise(average_AI = mean(AI2015), count = n(), 
            pct = count/count_all)
```

```{r}
hist(valid$AI)
hist(valid$AI_off)
hist(valid$AI2015)
```
CAL conversion for peaktime values 

```{r}
valid_CAL <- AI_and_PTAL%>%
  filter(AI > 0)
  
count_all_cal <- length(valid_CAL$AI)

cal_summary <- PTAL_summary%>%
  mutate(n_cal = pct*count_all_cal)

AI_and_PTAL<-AI_and_PTAL%>%
  arrange(AI)

b1a <- nth(AI_and_PTAL$AI, round(cal_summary$n_cal[2]), na_rm = TRUE)
b1b <- nth(AI_and_PTAL$AI, round(sum(cal_summary$n_cal[2:3])), na_rm = TRUE)
b2 <- nth(AI_and_PTAL$AI, round(sum(cal_summary$n_cal[2:4])), na_rm = TRUE)
b3 <- nth(AI_and_PTAL$AI, round(sum(cal_summary$n_cal[2:5])), na_rm = TRUE)
b4 <- nth(AI_and_PTAL$AI, round(sum(cal_summary$n_cal[2:6])), na_rm = TRUE)
b5 <- nth(AI_and_PTAL$AI, round(sum(cal_summary$n_cal[2:7])), na_rm = TRUE)
b6a <- nth(AI_and_PTAL$AI, round(sum(cal_summary$n_cal[2:8])), na_rm = TRUE)
b6b <- nth(AI_and_PTAL$AI, round(sum(cal_summary$n_cal[2:9])), na_rm = TRUE)

```

```{r}
CAL_and_PTAL_peak <- AI_and_PTAL%>%
  mutate(CAL = case_when(AI > 0 & AI <= b1a  ~ "1a",
                         AI > b1a & AI <= b1b  ~ "1b",
                         AI > b1b & AI <= b2   ~ "2",
                         AI > b2  & AI <= b3   ~ "3",
                         AI > b3  & AI <= b4   ~ "4",
                         AI > b4  & AI <= b5   ~ "5",
                         AI > b5  & AI <= b6a  ~ "6a",
                         AI > b6a   ~ "6b",
                         is.na(AI)~ "0",
                         TRUE ~ NA))
```

```{r}
count_all_cal <- length(CAL_and_PTAL_peak$CAL) - sum(is.na(CAL_and_PTAL_peak$CAL))

CAL_summary <- CAL_and_PTAL_peak%>%
  group_by(CAL)%>%
  summarise(average_AI = mean(AI), count = n(), 
            pct = count/count_all_cal)
```

CAL conversion for off-peak values 

```{r}
valid_CAL_off <- AI_and_PTAL%>%
  filter(AI_off > 0)
  
count_all_cal_off <- length(valid_CAL_off$AI_off)

cal_summary_off <- PTAL_summary%>%
  mutate(n_cal = pct*count_all_cal_off)

AI_and_PTAL<-AI_and_PTAL%>%
  arrange(AI_off)

b1a <- nth(AI_and_PTAL$AI_off, round(cal_summary_off$n_cal[2]), na_rm = TRUE)
b1b <- nth(AI_and_PTAL$AI_off, round(sum(cal_summary_off$n_cal[2:3])), na_rm = TRUE)
b2 <- nth(AI_and_PTAL$AI_off, round(sum(cal_summary_off$n_cal[2:4])), na_rm = TRUE)
b3 <- nth(AI_and_PTAL$AI_off, round(sum(cal_summary_off$n_cal[2:5])), na_rm = TRUE)
b4 <- nth(AI_and_PTAL$AI_off, round(sum(cal_summary_off$n_cal[2:6])), na_rm = TRUE)
b5 <- nth(AI_and_PTAL$AI_off, round(sum(cal_summary_off$n_cal[2:7])), na_rm = TRUE)
b6a <- nth(AI_and_PTAL$AI_off, round(sum(cal_summary_off$n_cal[2:8])), na_rm = TRUE)
b6b <- nth(AI_and_PTAL$AI_off, round(sum(cal_summary_off$n_cal[2:9])), na_rm = TRUE)

```

```{r}
CAL_and_PTAL_off <- AI_and_PTAL%>%
  mutate(CAL_off = case_when(AI_off > 0 & AI_off <= b1a  ~ "1a",
                         AI_off > b1a & AI_off <= b1b  ~ "1b",
                         AI_off > b1b & AI_off <= b2   ~ "2",
                         AI_off > b2  & AI_off <= b3   ~ "3",
                         AI_off > b3  & AI_off <= b4   ~ "4",
                         AI_off > b4  & AI_off <= b5   ~ "5",
                         AI_off > b5  & AI_off <= b6a  ~ "6a",
                         AI_off > b6a   ~ "6b",
                         is.na(AI_off)~ "0",
                         TRUE ~ NA))
```

```{r}
count_all_cal_off <- length(CAL_and_PTAL_off$CAL_off) - sum(is.na(CAL_and_PTAL_off$CAL_off))

CAL_summary_off <- CAL_and_PTAL_off%>%
  group_by(CAL_off)%>%
  summarise(average_AI = mean(AI_off), count = n(), 
            pct = count/count_all_cal_off)
```



## add grid polygon for mapping

```{r}
grid <- st_read(here::here("data/PTAL_grid_cell_2015/PTAL_100m_Grid_Cells.shp"))%>%
  st_transform(., 4326)%>%
  select("ID")

grid$ID <- grid$ID%>%
  as.character()

CAL_and_PTAL_off <-CAL_and_PTAL_off%>%
  st_drop_geometry()%>%
  select("from_id","CAL_off")

CAL_and_PTAL_peak <-CAL_and_PTAL_peak%>%
  st_drop_geometry()

CAL_and_PTAL_all <-CAL_and_PTAL_peak%>%
  left_join(.,CAL_and_PTAL_off,by= join_by("from_id"=="from_id"))
  
CAL_and_PTAL_all <-grid%>%
  left_join(.,CAL_and_PTAL_all,by= join_by("ID"=="from_id"))%>%
  st_as_sf()
```


```{r}
count_cal_off <- length(CAL_and_PTAL_all$CAL_off) - sum(is.na(CAL_and_PTAL_all$CAL_off))
count_cal_peak <- length(CAL_and_PTAL_all$CAL) - sum(is.na(CAL_and_PTAL_all$CAL))

CAL_band_off <- CAL_and_PTAL_all%>%
  st_drop_geometry()%>%
  group_by(CAL_off)%>%
  summarise(mini = min(AI_off), maxi = max(AI_off), frq = n()/count_cal_off)

CAL_band_peak <- CAL_and_PTAL_all%>%
  st_drop_geometry()%>%
  group_by(CAL)%>%
  summarise(mini = min(AI),maxi = max(AI),frq = n()/count_cal_peak)


```

```{r}
st_write(CAL_band_peak, "cal_band_peak.csv")
st_write(CAL_band_off, "cal_band_offpeak.csv")

```
## add london map for mapping 

```{r}
london <- st_read(here::here("data/London_Boroughs.gpkg"))%>%
  st_transform(., 4326)%>%
  mutate(inner_outer = case_when(ons_inner == "F" ~ "Outer London",
         TRUE ~ "Inner London"))
```

```{r}
tmap_mode("plot")

map_AI_cal =
  tm_shape(CAL_and_PTAL_all) +
  tm_polygons(col = "AI",palette="-viridis",
          n = 6,
          style = "quantile",
          scale = 3, 
          showNA = FALSE,
          colorNA = 'black',
          legend.hist = TRUE,
          title = "Access Index (CAL, peak)",
          border.alpha = 0) +
tm_shape(london) +
  tm_polygons(col = NA, 
              alpha = 0)+ 
tm_layout(title="AI",
             frame = FALSE,
             legend.outside = TRUE,
             legend.text.size = 1,
             legend.hist.size = 1,
             legend.hist.width = 0.75)

map_AI_cal

map_AI_cal_off =
  tm_shape(CAL_and_PTAL_all) +
  tm_polygons(col = "AI_off",palette="-viridis",
          n = 6,
          style = "quantile",
          scale = 3, 
          showNA = FALSE,
          colorNA = 'black',
          legend.hist = TRUE,
          title = "Access Index (CAL, off-peak)",
          border.alpha = 0) +
tm_shape(london) +
  tm_polygons(col = NA, 
              alpha = 0)+ 
tm_layout(title="AI",
             frame = FALSE,
             legend.outside = TRUE,
             legend.text.size = 1,
             legend.hist.size = 1,
             legend.hist.width = 0.75)

map_AI_cal_off

map_AI_ptal =
  tm_shape(CAL_and_PTAL_all) +
  tm_polygons(col = "AI2015",palette="-viridis",
          n = 6,
          style = "quantile",
          scale = 3, 
          showNA = FALSE,
          colorNA = 'black',
          legend.hist = TRUE,
          title = "Access Index (PTAL)",
          border.alpha = 0) +
tm_shape(london) +
  tm_polygons(col = NA, 
              alpha = 0)+ 
tm_layout(title="AI",
             frame = FALSE,
             legend.outside = TRUE,
             legend.text.size = 1,
             legend.hist.size = 1,
             legend.hist.width = 0.75)


map_AI_ptal
```

```{r}
tmap_save(map_AI_cal, 'london_ai_cal.png')
tmap_save(map_AI_cal_off, 'london_ai_off_cal.png')
tmap_save(map_AI_ptal, 'london_ai_ptal.png')

```

```{r}
hist(CAL_and_PTAL_all$AI)

```

```{r}
ggplot(data = CAL_and_PTAL_all, aes(x = CAL, fill=CAL)) +
    geom_bar(color="black")+
  theme_minimal()+
  theme(legend.position="none")+
  scale_fill_manual(values=c('white','#fde725',"#a0da39","#4ac16d", "#1fa187", "#277f8e", "#365c8d", "#46327e","#440154","black"))+
  ylim(0,45000)

ggplot(data = CAL_and_PTAL_all, aes(x = CAL_off, fill=CAL_off)) +
    geom_bar(color="black")+
  theme_minimal()+
    theme(legend.position="none")+
  scale_fill_manual(values=c('white','#fde725',"#a0da39","#4ac16d", "#1fa187", "#277f8e", "#365c8d", "#46327e","#440154","black"))+
  ylim(0,45000)

ggplot(data = CAL_and_PTAL_all, aes(x = PTAL2015, fill=PTAL2015)) +
    geom_bar(color="black")+
  theme_minimal()+
    theme(legend.position="none")+
  scale_fill_manual(values=c('white','#fde725',"#a0da39","#4ac16d", "#1fa187", "#277f8e", "#365c8d", "#46327e","#440154","black"))+
  ylim(0,45000)
```


```{r}
COLOR = c('white','#fde725',"#a0da39","#4ac16d", "#1fa187", "#277f8e", "#365c8d", "#46327e","#440154")
# this is just the Viridis colour code from https://waldyrious.net/viridis-palette-generator/ with white as the initial colour for NA/0 values
Labels = c('0: missing values', '1a: least accessible','1b','2','3','4','5','6a','6b: most accessible')

tmap_mode("plot")

map_cal =
  tm_shape(CAL_and_PTAL_all) +
  tm_polygons(col = "CAL",palette=COLOR,
          showNA = FALSE,
          colorNA = 'white',
          title = "CAL",
          border.alpha = 0,
          labels= Labels) +
tm_shape(london) +
  tm_polygons(col = NA, 
              alpha = 0)+ 
tm_layout(bg.color = NA, 
            legend.position = c(.85,0.05),
            main.title="Cycle Access Level (peak time)",
          main.title.size = 1,
            frame = FALSE,
           legend.text.size = 0.7)+
  tm_compass(position=c(0,0.1)) + 
  tm_scale_bar(breaks = c(0, 2, 4),position=c(0,0.03)) +
  tm_credits("Data source: OSM, GLA, TfL. R package: r5r.", fontface = "italic", position=c(0,0))

map_cal


map_cal_off =
  tm_shape(CAL_and_PTAL_all) +
  tm_polygons(col = "CAL_off",palette=COLOR,
          showNA = FALSE,
          colorNA = 'white',
          title = "CAL",
          border.alpha = 0,
          labels= Labels) +
tm_shape(london) +
  tm_polygons(col = NA, 
              alpha = 0)+ 
tm_layout(bg.color = NA, 
            legend.position = c(.85,0.05),
            main.title="Cycle Access Level (off-peak time)",
          main.title.size = 1,
            frame = FALSE,
           legend.text.size = 0.7)+
  tm_compass(position=c(0,0.1)) + 
  tm_scale_bar(breaks = c(0, 2, 4),position=c(0,0.03)) +
  tm_credits("Data source: OSM, GLA, TfL. R package: r5r.", fontface = "italic", position=c(0,0))

map_cal_off

map_ptal =
  tm_shape(CAL_and_PTAL_all) +
  tm_polygons(col = "PTAL2015",palette=COLOR,
          showNA = FALSE,
          colorNA = 'white',
          title = "PTAL",
          border.alpha = 0,
          labels= Labels) +
tm_shape(london) +
  tm_polygons(col = NA, 
              alpha = 0)+ 
tm_layout(bg.color = NA, 
            legend.position = c(.85,0.05),
            main.title="Public Transport Access Level",
          main.title.size = 1,
            frame = FALSE,
            legend.text.size = 0.7)+
  tm_compass(position=c(0,0.1)) + 
  tm_scale_bar(breaks = c(0, 2, 4),position=c(0,0.03)) +
  tm_credits("Data source: TfL.", fontface = "italic", position=c(0,0))

map_ptal
```

```{r}
tmap_save(map_cal, 'london_cal_peak.png')
tmap_save(map_cal_off, 'london_cal_offpeak.png')
tmap_save(map_ptal, 'london_ptal.png')
```

The above shows the Access Index values, before converted into the PTAL bands. Extremely high values tend to concentrate in inner London... Below shows the converted result using the frequency of PTAL bands, highlighting the most accessible locations.  

PTAL is related to time reliability of transport services, distance between OD pairs, number and frequency of transport services. However, the inner and outer London's difference in access levels is not represented completely in the PTAL map, but the AI map showing sharp contrast. This may mislead the interpretation of local connectivity in Outer London. The connectivity there can be better in some locations with tube stations, but in other parts, people may take longer to get to the transit. Housing development in those areas may still lead to high car dependency due to limited connectivity. 

## export the data into csv

```{r}
# drop Z dimension to export the data
new <- st_zm(CAL_and_PTAL_all, drop=T, what='ZM')

st_write(new, "london_cal.gpkg", driver = "GPKG",layer_options = "OVERWRITE=true")
```

# Spatial autocorrelation 

Areas with high accessibility can be also influential to their surrounding locations. People living in the grid close to high AI values can also enjoy the good proximity and accessibility through sustainable transport. Hence, spatial autocorrelation of the CAL or AI values will be assessed to see if there is any cluster of such places that could be the ideal locations for cycle infrastructure investment and new housing development projects.  

## Moran's I

- Spatial weight matrix 

```{r}
new_dropna <-new %>%
  filter(AI > 0)
```

```{r}
coordsW <- new_dropna%>%
  st_centroid()%>%
  st_geometry()%>%
  st_transform(.,27700)

new_27700 <- new_dropna%>%
  st_transform(.,27700)

nb <- new_27700%>%
  poly2nb(., queen=TRUE) #produce a neighbours list based on regions with contiguous boundaries

#plot them
plot(nb, st_geometry(coordsW), col="red")
```

```{r}
listw <- nb2listw(nb, zero.policy = TRUE)
```

- Computing the Global Moran’s I

```{r}
globalMoran <- moran.test(new_27700$AI, listw, zero.policy=TRUE,na.action=na.omit)
globalMoran

I <- globalMoran[["estimate"]][["Moran I statistic"]]
I
pvalue <- globalMoran[["p.value"]]
pvalue
```

Global morans I statistic is 0.9368 with p value close to 0. The p-value is statistically significant, and the z-score is positive. This reveals that the Access index is not randomly distributed among the features in Camden and the AI is clustered with similar values. 

To see the location of the hot spots, local morans i can be measured for each grid. This project includes codes that are adapted from https://rpubs.com/quarcs-lab/spatial-autocorrelation. 

Local spatial autocorrelation

```{r}
moran <- moran.plot(new_27700$AI, listw = nb2listw(nb, style = "W", zero.policy = TRUE), zero.policy = TRUE, na.action= na.omit)
```
- Compute local Moran

```{r}
local <- localmoran(x = new_27700$AI, listw = nb2listw(nb, style = "W", zero.policy = TRUE), zero.policy = TRUE, na.action= na.omit)

moran.map <- cbind(new_27700, local)

tm_shape(moran.map) +
  tm_fill(col = "Ii",
          style = "quantile",
          title = "local moran statistic") +
  tm_layout(title="local morans I",
             frame = FALSE,
             legend.outside = TRUE,
            legend.text.size = 1,
             legend.hist.size = 1,
             legend.hist.width = 0.75)
```


- Plot LISA clusters

```{r}
quadrant <- vector(mode="numeric",length=nrow(local))

# centers the variable of interest around its mean
m.AI <- new_27700$AI - mean(new_27700$AI)     

# centers the local Moran's around the mean
m.local <- local[,1] - mean(local[,1])    

# significance threshold
signif <- 0.1 

# builds a data quadrant
quadrant[m.AI >0 & m.local>0] <- "4"  
quadrant[m.AI <0 & m.local<0] <- "1"      
quadrant[m.AI <0 & m.local>0] <- "2"
quadrant[m.AI >0 & m.local<0] <- "3"
quadrant[local[,5]>signif] <- "0"   


quadrant.map <- cbind(new_27700, quadrant)

quadrant.map <- quadrant.map %>%
  mutate(cluster = case_when(quadrant == "0" ~ "insignificant",
                             quadrant == "1" ~ "low-low",
                             quadrant == "2" ~ "low-high",
                             quadrant == "3" ~ "high-low",
                             quadrant == "4" ~ "high-high"))

cluster.map <- quadrant.map%>%
  select("geometry","cluster")
```

```{r}
brks <- c("0","1","2","3","4")
colors <- c("white","blue",rgb(0,0,1,alpha=0.4),rgb(1,0,0,alpha=0.4),"red")
Palette <- c("white", "blue", "lightblue", "pink", "red" )

Labels = c("insignificant", "low-low", "low-high", "high-low" ,"high-high")
  
tmap_mode("plot")

map_cluster = tm_shape(quadrant.map) +
  tm_fill(col="quadrant",
          breaks = brks,
          palette=Palette,
          showNA = FALSE,
          colorNA = 'white',
          title = "cluster",
          border.alpha = 0,
          labels= Labels) +
tm_shape(london) +
  tm_polygons(col = NA, 
              alpha = 0)+ 
  tm_layout(bg.color = NA, 
            legend.position = c(.85,0.05),
            main.title="LISA cluster map",
            main.title.size = 1,
             frame = FALSE,
             legend.text.size = 0.7)+
  tm_compass(position=c(0,0.1)) + 
  tm_scale_bar(breaks = c(0, 2, 4),position=c(0,0.03)) +
  tm_credits("Data source: OSM, GLA, TfL. R package: r5r.", fontface = "italic", position=c(0,0))

map_cluster
```

This output shows that inner London as a large cluster generally has high CAL surrounded by also high CAL. Outer London on the other hand has lower CAL and surrounded by low CAL grids. Good local connectivity is concentrated in central locations while outer boroughs obtain poorer local connectivity.Those areas are more likely to be car dependent as the local connectivity through walking, cycling and public transit is poor. The middle ring between inner and outer London clusters shows not significant result. A mixture of high and low CAL is present in this ring, implying high potential for planners to direct less infrastructure into the areas to gain higher return in improving local connectivity. 


```{r}
tmap_save(map_cluster, 'LISA_cluster.png')
```

# CAL per borough

Average CAL will then be measured per borough to identify case study area in this project. And compare between the inner and outer london boroughs. 

Add borough info to each grid. 

```{r}
CAL_and_PTAL_borough_grid <-  st_join(CAL_and_PTAL_all, london)
```

```{r}
CAL_and_PTAL_borough_grid <- CAL_and_PTAL_borough_grid%>%
  unique()
```

```{r}
CAL_and_PTAL_borough <- CAL_and_PTAL_borough_grid %>%
  mutate(AI_mean = (AI+AI_off)/2)%>%
  group_by(name)%>%
  summarize(average_AI=mean(AI_mean,na.rm=TRUE), inner_outer=first(inner_outer))
```

```{r}
brent <- london%>%
  filter(name == "Brent")
brom <- london%>%
  filter(name == "Bromley")
```

```{r}
tmap_mode("plot")

map_mean_ai = tm_shape(CAL_and_PTAL_borough) +
  tm_fill(col="average_AI",
          palette="-viridis",
          title = "average AI",
          border.alpha = 0) +
tm_shape(london) +
  tm_polygons(col = NA, 
              alpha = 0)+ 
tm_shape(brent) +
  tm_polygons(col = NA, 
              border.col = "red",
              alpha = 0)+ 
  tm_text( "name")+
tm_shape(brom) +
  tm_polygons(col = NA, 
              border.col = "red",
              alpha = 0)+ 
  tm_text( "name")+
  tm_layout(bg.color = NA, 
            legend.position = c(.85,0.05),
            main.title="Average Access Index map",
            main.title.size = 1,
             frame = FALSE,
             legend.text.size = 0.7)+
  tm_compass(position=c(0,0.1)) + 
  tm_scale_bar(breaks = c(0, 2, 4),position=c(0,0.03)) +
  tm_credits("Data source: OSM, GLA, TfL. R package: r5r.", fontface = "italic", position=c(0,0))

map_mean_ai
```
```{r}
tmap_save(map_mean_ai, 'mean_ai_borough.png')
```

```{r}
mean_ai_inner_outer <- CAL_and_PTAL_borough %>%
  group_by(inner_outer) %>%
  summarise(mean=mean(average_AI, na.rm=TRUE))

mean_ai_inner_outer
```

```{r}
library(ggplot2)
library(forcats)

bar <- ggplot(CAL_and_PTAL_borough, aes(x = fct_reorder(name, desc(average_AI)), y = average_AI, fill = inner_outer)) + 
   geom_bar(stat = "identity") + 
   labs(title = "Average AI in London boroughs by inner and outer boroughs",
        x = "Borough name", y = "Mean AI") +
    theme_classic()+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +          # Rotate axis labels
    coord_fixed(1/30)
bar
```


```{r}
png("bar_inner_outer.png")
print(bar)
dev.off()
```

# Objective 3: identify areas of potential using data from PTAL and CAL 



```{r}
CAL_and_PTAL_brent <- CAL_and_PTAL_borough_grid%>%
  filter(name == "Brent")
CAL_and_PTAL_brom <- CAL_and_PTAL_borough_grid%>%
  filter(name == "Bromley")
```

```{r}
COLOR = c('white','#fde725',"#a0da39","#4ac16d", "#1fa187", "#277f8e", "#365c8d", "#46327e","#440154")
COLORcal = c('white',"#a0da39","#4ac16d", "#1fa187", "#277f8e", "#365c8d", "#46327e","#440154")

tmap_mode("plot")

map_cal_brent =
  tm_shape(CAL_and_PTAL_brent) +
  tm_polygons(col = "CAL",palette=COLORcal,
          showNA = FALSE,
          colorNA = 'white',
          border.alpha = 0,
          title="") +
  tm_shape(brent) +
  tm_polygons(col = NA, 
              alpha = 0)+ 
tm_layout(bg.color = NA, 
          panel.labels = "Cycle Access Level",
          panel.label.size = 0.8,
          panel.label.bg.color = NA,
          frame = FALSE,
          legend.text.size = 0.7,
          legend.position = c("right","top"))+
  tm_compass(position=c(0,0.1)) + 
  tm_scale_bar(breaks = c(0, 2, 4),position=c(0,0.03)) +
  tm_credits("Data source: OSM, GLA, TfL. R package: r5r.", fontface = "italic", position=c(0,0))

map_ptal_brent =
  tm_shape(CAL_and_PTAL_brent) +
  tm_polygons(col = "PTAL2015",palette=COLOR,
          showNA = FALSE,
          colorNA = 'white',
          border.alpha = 0,
          title="") +
  tm_shape(brent) +
  tm_polygons(col = NA, 
              alpha = 0)+ 
tm_layout(bg.color = NA, 
          panel.labels = "Public Transport Access Level",
          panel.label.size = 0.8,
          panel.label.bg.color = NA,
          frame = FALSE,
          legend.text.size = 0.7,
          legend.position = c("right","top"))+
  tm_credits("Data source: TfL.", fontface = "italic", position=c(0,0))


t <- tmap_arrange(map_cal_brent, map_ptal_brent, ncol =2,asp = 1)
t
```
```{r}
tmap_save(t, 'brent_cal_ptal.png')
```



```{r}
tmap_mode("plot")

map_cal_brom =
  tm_shape(CAL_and_PTAL_brom) +
  tm_polygons(col = "CAL",palette=COLOR,
          showNA = FALSE,
          colorNA = 'white',
          border.alpha = 0,
          title="") +
  tm_shape(brom) +
  tm_polygons(col = NA, 
              alpha = 0)+ 
tm_layout(bg.color = NA, 
          panel.labels = "Cycle Access Level",
          panel.label.size = 0.8,
          panel.label.bg.color = NA,
            frame = FALSE,
           legend.text.size = 0.7,
            legend.position = c("right","bottom"))+
  tm_compass(position=c(0,0.1)) + 
  tm_scale_bar(breaks = c(0, 2, 4),position=c(0,0.03)) +
  tm_credits("Data source: OSM, GLA, TfL. R package: r5r.", fontface = "italic", position=c(0,0))

map_ptal_brom =
  tm_shape(CAL_and_PTAL_brom) +
  tm_polygons(col = "PTAL2015",palette=COLOR,
          showNA = FALSE,
          colorNA = 'white',
          border.alpha = 0,
          title="") +
  tm_shape(brom) +
  tm_polygons(col = NA, 
              alpha = 0)+ 
tm_layout(bg.color = NA, 
          panel.labels = "Public Transport Access Level",
          panel.label.size = 0.8,
          panel.label.bg.color = NA,
            frame = FALSE,
            legend.text.size = 0.7,
          legend.position = c("right","bottom"))+
  tm_credits("Data source: TfL.", fontface = "italic", position=c(0,0))


t2 <- tmap_arrange(map_cal_brom, map_ptal_brom, ncol =2,asp = 1)
t2
```
```{r}
tmap_save(t2, 'bromley_cal_ptal.png')
```


#  bivariate choropleth maps

```{r}
# install.packages("biscale")
# install.packages("cowplot")
# install.packages("ggsn")
library(biscale)
library(cowplot)
library(ggsn)

```

```{r}

bi_cho_data <- function (input) {
  
  input2<-input%>%
    mutate(cal_num = case_when(CAL == "1a" | CAL == "1b" | CAL == "2"~ 1, 
                             CAL == "3" | CAL == "4"~ 2,
                             CAL == "5" | CAL == "6a" | CAL == "6b"~ 3,
                             TRUE ~ NA))%>%
    mutate(ptal_num = case_when(PTAL2015 == "1a" | PTAL2015 == "1b" | PTAL2015 == "2"~ 1, 
                             PTAL2015 == "3" | PTAL2015 == "4"~ 2,
                             PTAL2015 == "5" | PTAL2015 == "6a" | PTAL2015 == "6b"~ 3,
                             TRUE ~ NA))%>%
    filter(!is.na(cal_num) & !is.na(ptal_num))
  
  data <- bi_class(input2, x = cal_num, y = ptal_num, style = "quantile", dim = 3)

  # create map

  map <- ggplot() +
  geom_sf(data = data, mapping = aes(fill = bi_class), color = NA, size = 0.1, show.legend = FALSE) +
    bi_scale_fill(pal = "GrPink", dim = 3) +
    labs(title = "Bivariate choropleth map (CAL and PTAL)") +
    bi_theme()+
    theme(plot.title = element_text(size=8))
  
  # county outline
  
  legend <- bi_legend(pal = "GrPink",
                    dim = 3,
                    xlab = "Higher CAL",
                    ylab = "Higher PTAL",
                    size = 4)
  finalPlot <- ggdraw() +
    draw_plot(map, 0, 0, 1, 1) +
    draw_plot(legend, .74, .74, .2, .2)
  
  
  return(finalPlot)
}
  
bi_cho_data(CAL_and_PTAL_brent)  
```


```{r}
ggsave("bi_cho_brent.png", width = 13, height = 13, units = "cm")
```

```{r}
bi_cho_data(CAL_and_PTAL_brom)  

ggsave("bi_cho_brom.png",width = 13, height = 13, units = "cm")
```
