
# MCDA

This section would be conducted after the 3_london_CAL_cont.Rmd file which identifies the case study borough of Brent and Bromley. This section aims to produce tilted stacked maps (so not MCDA technically) of various attributes to give a context for the CAL result. 

```{r}
library(osmextract)
library(elevatr)
library(raster)
library(stars)
library(CycleInfraLnd)
library(ggnewscale)
```

The stacked maps include the following attributes for both boroughs: 
- elevation
- population density per LSOA
- cycle parking capacity and location  
- road network

# load data 

```{r}

ldn <-  read_sf("data/London_Boroughs.gpkg")%>%
    st_transform(., 4326)

brent <- ldn%>%
  filter(name == "Brent")

brom <- ldn %>%
  filter(name =="Bromley")

# elevation

dem <- stars::read_stars(file.path("C:/Users/99/Documents/CASA/dissertation_test/input/london_elevation.tif"), package = "stars")
dem <- st_as_sf(dem)

dem1 <- dem[brent,]
dem2 <- dem[brom,]

dem1 <- dem1%>%
  rename("elevation" = "london_elevation.tif")
dem2 <- dem2%>%
  rename("elevation" = "london_elevation.tif")
  
# road network (OSM)

roads <- oe_read(
 file.path("C:/Users/99/Documents/CASA/dissertation_test/input/geofabrik_greater-london-latest.osm.pbf"), layer = "lines")

# crop
roads1 <- roads[brent,]
roads1 <- st_crop(roads1, brent)
plot(roads1['osm_id'])

roads2 <- roads[brom,]
roads2 <- st_crop(roads2, brom)
plot(roads2['osm_id'])

roads1 <- roads1 %>%
  filter(highway == "primary" |highway == "secondary"| highway == "tertiary"| highway == "cycleway")

roads2 <- roads2 %>%
  filter(highway == "primary" |highway == "secondary"| highway == "tertiary"| highway == "cycleway")

# cycle data 


cid_parking <- get_cid_points(type = "cycle_parking")

cid_parking_brent <- cid_parking[brent,]%>%
  dplyr::select("FEATURE_ID","PRK_CPT","BOROUGH","geometry")%>%
  mutate(capacity = case_when(PRK_CPT < 10 ~ "1-10",
                              PRK_CPT >= 10 & PRK_CPT <20 ~ "10-20",
                              PRK_CPT >= 20 & PRK_CPT <30 ~ "20-30",
                              PRK_CPT >=30 ~ "30+"))


cid_parking_brom <- cid_parking[brom,]%>%
  dplyr::select("FEATURE_ID","PRK_CPT","BOROUGH","geometry")%>%
  mutate(capacity = case_when(PRK_CPT < 10 ~ "1-10",
                              PRK_CPT >= 10 & PRK_CPT <20 ~ "10-20",
                              PRK_CPT >= 20 & PRK_CPT <30 ~ "20-30",
                              PRK_CPT >=30 ~ "30+"))

# population 


pop_brent <- read.csv(here::here("data/brent_popden_2021lsoa.csv"), 
                        sep = ",",  header = TRUE, 
                         encoding = "latin1")%>%
    dplyr::select("LSOA21code","pop_den")

  
pop_brom <- read.csv(here::here("data/bromley_popden_2021lsoa.csv"), 
                        sep = ",",  header = TRUE, 
                         encoding = "latin1")%>%
  dplyr::select("LSOA21code","pop_den")

# read in lsoa 2011 code   
lsoa <- st_read(here::here("data/lsoa_geo_shp/LSOA_2011_London_gen_MHW.shp"))

lsoa <- lsoa %>%
  st_transform(4236)%>%
  dplyr::select("LSOA11CD","LSOA11NM","LAD11CD","LAD11NM", "geometry")

# find changes in code from 2011 to 2021
lsoa22 <- read.csv(here::here("data/lsoa11to22.csv"),sep = ",",  header = TRUE, 
                         encoding = "latin1")
lsoa22 <- lsoa22%>%
  filter(LAD22NM =="Brent" | LAD22NM == "Bromley")%>%
  filter(LSOA11CD != LSOA21CD)%>%
  dplyr::select("LSOA11CD", "LSOA21CD")

lsoa_brent <- lsoa%>%
  filter(LAD11NM == "Brent")%>%
  left_join(., lsoa22, by = join_by("LSOA11CD" == "LSOA11CD"))%>%
  mutate(code = case_when(!is.na(LSOA21CD) ~ LSOA21CD,
                          TRUE~ LSOA11CD))
lsoa_bromley <- lsoa%>%
  filter(LAD11NM == "Bromley")%>%
  left_join(., lsoa22, by = join_by("LSOA11CD" == "LSOA11CD"))%>%
  mutate(code = case_when(!is.na(LSOA21CD) ~ LSOA21CD,
                          TRUE~ LSOA11CD))

#join by new code
pop_brent <- lsoa_brent%>%
  left_join(.,pop_brent, by = join_by( "code" == "LSOA21code"))
pop_brom <- lsoa_bromley%>%
  left_join(.,pop_brom, by = join_by("code" == "LSOA21code"))

pop_brent<-pop_brent%>%
  dplyr::select("code", "pop_den","geometry")
pop_brom<-pop_brom%>%
  dplyr::select("code", "pop_den","geometry")

pop_brent <- pop_brent%>%
  rename("population_density"= "pop_den")
pop_brom <- pop_brom%>%
  rename("population_density"= "pop_den")

plot(pop_brent)
plot(pop_brom)

```
This tilted stacked map is produced with credit to the code here: https://www.urbandemographics.org/post/figures-map-layers-r/

Apart from using the function to rotate the graphics, a package can also be used: https://github.com/marcosci/layer


```{r}
rotate_data <- function(data, x_add = 0, y_add = 0) {
  
  shear_matrix <- function(){ matrix(c(2, 1.2, 0, 1), 2, 2) }
  
  rotate_matrix <- function(x){ 
    matrix(c(cos(x), sin(x), -sin(x), cos(x)), 2, 2) 
  }
  data %>% 
    dplyr::mutate(
      geometry = .$geometry * shear_matrix() * rotate_matrix(pi/20) + c(x_add, y_add)
    )
}

rotate_data_geom <- function(data, x_add = 0, y_add = 0) {
  shear_matrix <- function(){ matrix(c(2, 1.2, 0, 1), 2, 2) }
  
  rotate_matrix <- function(x) { 
    matrix(c(cos(x), sin(x), -sin(x), cos(x)), 2, 2) 
  }
  data %>% 
    dplyr::mutate(
      geom = .$geom * shear_matrix() * rotate_matrix(pi/20) + c(x_add, y_add)
    )
}
```

```{r}
### plot 

# annotate parameters
x = 68.43
color = 'gray40'

temp1 <- ggplot() +
          
        # terrain
        geom_sf(data = dem1 %>% rotate_data(), aes(fill=elevation), color=NA) +
        scale_fill_distiller(palette = "YlOrRd", direction = 1) +
        annotate("text", label='Elevation', x=x, y= 41.4, hjust = 0, color=color) +
        labs(caption = "Data source: ONS, OSM, TfL. R package: elevatr.")

temp2 <- temp1 +
        
        # pop
        new_scale_fill() + 
        new_scale_color() +
        geom_sf(data = pop_brent %>% rotate_data(y_add = .15), aes(fill=population_density), color=NA) +
        scale_fill_viridis_c(option = "inferno", direction = -1) +
        annotate("text", label='Population density by LSOA (2021)', x=x, y= 41.55, hjust = 0, color=color) + 

        # osm
        new_scale_fill() + 
        new_scale_color() +
        geom_sf(data = roads1 %>% rotate_data(y_add = .3), aes(color=highway), size=.2) +
        scale_colour_viridis_d(option = "viridis") +
        annotate("text", label='Road hierarchy', x=x, y= 41.7, hjust = 0, color=color) +
        
        # cycle parking
        new_scale_fill() + 
        new_scale_color() +
        geom_sf(data = cid_parking_brent %>% rotate_data(y_add = .45), aes(color=capacity), size = .5) +
        scale_colour_viridis_d(option = "mako", begin = 0, end = 0.8, direction = -1) +
        annotate("text", label='Cycle parking capacity', x=x, y= 41.85, hjust = 0, color=color)+
      
        theme_void()
temp2
        
```

```{r}
ggsave(plot = temp2, filename = 'map_layers_brent.png', 
       dpi=200, width = 15, height = 16, units='cm')
```

```{r}
### plot  ----------------

# annotate parameters
x = 68.7
color = 'gray40'

temp3 <- ggplot() +
          
        # terrain
        geom_sf(data = dem2 %>% rotate_data(), aes(fill=elevation), color=NA) +
        scale_fill_distiller(palette = "YlOrRd", direction = 1) +
        annotate("text", label='Elevation', x=x, y= 41.2, hjust = 0, color=color) +
        labs(caption = "Data source: ONS, OSM, TfL. R package: elevatr.")

temp4 <- temp3 +
        
        # pop
        new_scale_fill() + 
        new_scale_color() +
        geom_sf(data = pop_brom %>% rotate_data(y_add = .2), aes(fill=population_density), color=NA) +
        scale_fill_viridis_c(option = "inferno", direction = -1) +
        annotate("text", label='Population density by LSOA (2021)', x=x, y= 41.4, hjust = 0, color=color) + 

        # osm
        new_scale_fill() + 
        new_scale_color() +
        geom_sf(data = roads2 %>% rotate_data(y_add = .4), aes(color=highway), size=.2) +
        scale_colour_viridis_d(option = "viridis") +
        annotate("text", label='Road hierarchy', x=x, y= 41.6, hjust = 0, color=color) +
        
        # cycle parking
        new_scale_fill() + 
        new_scale_color() +
        geom_sf(data = cid_parking_brom %>% rotate_data(y_add = .6), aes(color=capacity), size = .5) +
        scale_colour_viridis_d(option = "mako", begin = 0, end = 0.8, direction = -1) +
        annotate("text", label='Cycle parking capacity', x=x, y= 41.8, hjust = 0, color=color)+
        theme_void()
temp4
   

       
```

```{r}
ggsave(plot = temp4, filename = 'map_layers_brom.png', 
       dpi=200, width = 15, height = 16, units='cm')
```

