---
title: "cycle_infra"
output: html_document
date: "2023-06-11"
---

This section aims to get an understanding of the cycle infrastructure database (CID) by TfL, more specifically the cycle parking AND cycle lanes/tracks data in London. It will start from using a sample of London's data - data in Camden borough to test and map the location of those cycling infrastructure and their attributes (e.g. capacity of cycle parking space). 

The CID data is in the form of R package, so this will be installed first. 

```{r}
#library(devtools)
#devtools::install_github("PublicHealthDataGeek/CycleInfraLnd")
```

# Load packages

```{r}
library(dplyr)
library(tmap)
library(sf)
```

```{r}
library(CycleInfraLnd)

# CID has either point or line data. 

?get_cid_points
?get_cid_lines
```

# Read in data from CID 

For point data, there are five types of infra - 

1) Signals - early-release signals at junctions
2) Cycle parking, including the type and capacity of parking   
3) Restricted Points – points where people cycling will have to dismount Paths through parks and other green spaces that can, and cannot, be cycled on
4) Signage - Signed cycle routes and other wayfinding
5) Traffic calming, including the location of all speed humps in Greater London

For this project, only cycle parking is in need. 

```{r}
cid_parking = get_cid_points(type = "cycle_parking")
```

For point data, there are five types of infra - 

1) Advanced stop lines – boxes at junctions for people cycling
2) Signalised crossings for cycles
3) Cycle lanes and tracks – including whether they are segregated or painted lanes
4) Restricted route - Modal filters and traffic gates which allow cycles to pass but restrict car traffic

For this project, Cycle lanes and tracks and Restricted route can be read in. 

```{r}
cycle_lane = get_cid_lines(type = "cycle_lane_track")
cycle_restricted = get_cid_lines(type = "restricted_route")
```

# Read in camden boundary

```{r}
London_borough <- st_read(here::here("data/London_Boroughs.gpkg"))%>%
    st_transform(., 4326)

camden = London_borough %>% 
  filter(name == "Camden")
```

# spatial subsetting to reduce data for mapping 

```{r}
cid_parking_cam = cid_parking[camden, , op = sf::st_within]
cycle_lane_cam = cycle_lane[camden, , op = sf::st_within]
cycle_restricted_cam = cycle_restricted[camden, , op = sf::st_within]
```

# map the infrastructure 

```{r}
tmap_mode("plot")
tm_shape(cid_parking_cam) +
  tm_dots(col="PRK_CPT",palette="RdYlBu")+
tm_shape(camden)+
  tm_polygons(col=NA,alpha=0)
  
tm_shape(cycle_lane_cam) +
  tm_lines(col="CLT_SEGREG",palette = "RdYlBu")+
tm_shape(camden)+
  tm_polygons(col=NA,alpha=0)

tmap_mode("plot")
tm_shape(cycle_lane_cam) +
  tm_lines(col="blue")+
tm_shape(cycle_restricted_cam) +
  tm_lines(col="black")+
tm_shape(camden)+
  tm_polygons(col=NA,alpha=0)
```

# compare the cycle tracks/lanes with OSM network data

```{r}
# Indicate the path where OSM and GTFS data are stored
r5r_core <- setup_r5(data_path = "C:/Users/99/Documents/CASA/dissertation_test/input")
```

```{r}
street_net <- street_network_to_sf(r5r_core)
street_net_l <- street_net$edges
street_net_cam = street_net_l[camden, , op = sf::st_within]
```

```{r}
tmap_mode("plot")

tm_shape(street_net_cam) +
  tm_lines(col="grey")+
tm_shape(cycle_lane_cam) +
  tm_lines(col="blue")+
tm_shape(cycle_restricted_cam) +
  tm_lines(col="black")+
tm_shape(camden)+
  tm_polygons(col=NA,alpha=0)
```


